---
title: "04-Model Formatting"
output: html_document
---
## Library requirements

```{r results='hide',message=FALSE}
if (!require("R.utils")) install.packages("R.utils")
library("R.utils")
if (!require("data.table")) install.packages("data.table")
library("data.table")
if (!require("googledrive")) install.packages("googledrive")
library("googledrive")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("foreach")) install.packages("foreach")
library(foreach)
if (!require("doSNOW")) install.packages("doSNOW")
library(doSNOW)
if (!require("caret")) install.packages("caret")
library(caret)
if (!require("mltools")) install.packages("mltools")
library(mltools)
if (!require("mlr")) install.packages("mlr")
library("mlr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("assertr")) install.packages("assert")
library("assertr")
```

### One-hot encoding
First, we read in the sample data, ignoring the indexing column.
```{r}
random_sample1 = fread('../data/processed/labelled.csv', colClasses=list(character=2))[,- 1]
```

We now look at each column to see if their typing is correct and if the column can be removed because they either do not have value for the analysis (uid) or only have one value (local_orig).
```{r}
random_sample1 <- random_sample1[, ':=' (uid=NULL,duration=as.numeric(duration),orig_bytes=as.numeric(orig_bytes),resp_bytes=as.numeric(resp_bytes),local_orig=NULL,tunnel_parents=NULL,missed_bytes=NULL)]
```
We now check for the categorical variable, which are those with type of character but are not the columns ts, srcIP and dstIP. Ideally, we would have wanted to one-hot encode the IP addresses as well, however this makes the table too big to store in R so we chose a different method, which will be seen later.
```{r}

categ={}
categoricals <- random_sample1[,sapply(.SD,is.character)]
for (i in 1:length(categoricals)){
  if (categoricals[i]){
    categ <- c(categ,i)
  }
}
categ
#we remove atcktype
categ <- categ[-length(categ)]
#remove ts,srcIP,dstIP
categ <- categ[- 1]
categ <- categ[- 1]
categ <- categ[- 1]
```

We now create a table with columns with the one-hot encoding of the categorical variables.
```{r}
dummy <- dummyVars(" ~ .",data=random_sample1[,.SD,.SDcols=c(categ)])
random_sample <- data.table(predict(dummy,newdata = random_sample1[,.SD,.SDcols=c(categ)]))
```

We then merged the one-hot encoding with the main sample table.
```{r}
random_sample$id = 1:nrow(random_sample)
random_sample1$id = 1:nrow(random_sample1)
random_sample1 = merge(random_sample1,random_sample,by.x='id',by.y = 'id') [,- 1]
random_sample1 = random_sample1[,-c('proto','service','conn_state','history','atcktype')]
```

Here we set all the Nas to 0.
```{r}
random_sample1$duration[is.na(random_sample1$duration)] = 0
random_sample1$orig_bytes[is.na(random_sample1$orig_bytes)] = 0
random_sample1$resp_bytes[is.na(random_sample1$resp_bytes)] = 0
```

---the above is just copied from the one-hot encoding in the previous section for the sake of consistency/continuity---

## IP Encoding
As mentioned above, we take a different approach to encoding the IP addresses instead of one-hot encoding. Having explored the following paper, https://doi.org/10.25394/PGS.11307287.v1 , we follow the binary encoding method suggested in the paper. Ideally, we would have followed the seemingly better performing split method wherein we split IPv4 addresses into 4 separate features e.g. 'IP': 192.168.202.24 would be split to features 'IP1':192 , 'IP2': 168 , ... and so on. However, the presence of IPv6 addresses in our dataset is problematic for this approach and so we instead use the binary split method highlighted below.


We start by creating a data table comprised only of the IP addresses.
```{r}
ip_df <- data.table(random_sample1[,c('srcIP','dstIP')])
```

Starting with srcIPs we find all instances of IPv6 addresses
```{r}
src_ipv6<-vector()

for (ip in ip_df[,srcIP]){
  has_colon <- grepl(':',ip)
  
  if (has_colon){
    src_ipv6<-append(src_ipv6,ip)
  }
  
}

head(src_ipv6)
```

Now, we find all the unique instaces of our IPv6 addresses so we can study their form.
```{r}
u_src_ipv6=unique(src_ipv6)
head(u_src_ipv6)
```
Of all the unique IPv6 addresses, we extract the ones that have the '::' short-hand in them so that we can later conver them to the full-form a1:a2:a3:...:a8
```{r}
src_double_colon<-vector()

for (u_ip in u_src_ipv6){
  if (grepl('::',u_ip)){
    src_double_colon<-append(src_double_colon,u_ip)
  }
}

src_double_colon
```

For somewhat better readability of src_double_colon's form. Uncomment if you want to inspect.
```{r}
# for (i in src_double_colon){
#   print(strsplit(i,'::'))
# }
```

For the short-hand IPv6 addresses found in srcIP, it looks like the form followed is either exactly '::' or 'fe80::ak:a(k+1):...:a8' where the value of k seems to vary. 


We now repeat the process for dstIP and see if there is any difference in the form of the IPv6 addresses.
```{r}
dst_ipv6<-vector()

for (ip in ip_df[,dstIP]){
  has_colon <- grepl(':',ip)
  
  if (has_colon){
    dst_ipv6<-append(dst_ipv6,ip)
  }
  
}

u_dst_ipv6=unique(dst_ipv6)

dst_double_colon<-vector()

for (u_ip in u_dst_ipv6){
  if (grepl('::',u_ip)){
    dst_double_colon<-append(dst_double_colon,u_ip)
  }
}

dst_double_colon
```
Again, for somewhat better readability of dst_double_colon's form. Uncomment to inspect.
```{r}
# for (i in dst_double_colon){
#   print(strsplit(i,'::'))
# }
```

For the dstIPs we have the same case of '::' and 'fe80::x' as before, but we also have a new form of '2001:a2:a3:a4::a8' where obviously a5==a6==a7==0000. We thus create a function to format all IPs as below.
```{r}
cleanIP <- function(ip) {
  ip<-as.character(ip)
  if(grepl('::',ip)){
    #special case
    if(ip=='::'){
      
      ip<-'0000:0000:0000:0000:0000:0000:0000:0000'
      
    }else {
      
      #all our IPv6 instances that have '::' within them are of the form
      #fe80::x, where x is some a1:a2:...:ak for k=[1:6]
      ip_split<-as.list(strsplit(ip,'::')[[1]])
      
      #this corresponds to a1:a2:...:ak mentioned above
      ip_sec_half<-ip_split[[2]]
      
      #ipv6 that start with 2 and contain :: follow the form 
      #a1:a2:a3:a4::a8, so 5-7 are 'missing'
      if (substring(ip,1,1)=='2'){
        
        missing_colons=3
        
        } else {
          #returns k mentioned above
          active_colons<-length(as.list(strsplit(ip_sec_half,':'))[[1]])
          #there are a total of 8 'sections' in an IPv6 address
          #including 'k' and 'fe80' we have m missing_colons below
          missing_colons=8-(active_colons+1)
        }
      
      #returns ':0000:0000:0000:' for m=3 for example
      zeros<-paste(c(replicate(missing_colons,':0000'),':'),collapse="")
      
      #putting all components together
      ip_clean<-paste(ip_split[[1]],zeros,ip_split[[2]],sep="")
      
      ip<-ip_clean
    }

  }
  return(ip)
}
```

We use the function below:
```{r}
#test_ip is just backup for ip_df 
test_ip<-data.table(ip_df)
#length of dataset
L = length(test_ip[,dstIP])
```

We now create a vector of indices for where IPv6 addresses appear
```{r}
src_ind = vector()
```

We initially did a separate list for dst_ind but then realised the indices are exactly the same so we just kept src_ind and got rid of `dst_ind = vector()` to reduce compute time.
```{r}
#progress bar
pb = txtProgressBar(min = 0, max = L, initial = 0)

#loop over to append i to src_ind
####this is really slow, not sure if there's a faster way to do this that would save time
for (i in 1:L){
  #ith entry of srcIP and dstIP
  sip=test_ip[,'srcIP'][i]
  dip=test_ip[,'dstIP'][i]

  if (grepl(':',sip)){
    
    sip=as.character(sip)
    dip=as.character(dip)
    #clean values
    sip=cleanIP(sip)
    dip=cleanIP(dip)
    #assign values
    test_ip[,'srcIP'][i]=sip
    test_ip[,'dstIP'][i]=dip
    #append our ipv6 index
    src_ind<-append(src_ind,i)
  }
  
  #progress step
  setTxtProgressBar(pb,i)
  
}
```

Quick check to see if values have changed in test_ip[,srcIP] or not
```{r}
suip = unique(test_ip[,srcIP])
duip = unique(test_ip[,dstIP])

print('----srcIP----')
suip6<-vector()
for (ip in suip){
  
  if (grepl('::',ip)){
    print(ip)
    suip6<-append(suip6,ip)
  }
}

print('----dstIP----')
duip6<-vector()
for (ip in duip){
  
  if (grepl('::',ip)){
    print(ip)
    duip6<-append(duip6,ip)
  }
}
```
Everything looks like it works fine. To summarise the method that we have been following and will be applying below:
 - we create a function that takes the 4 elements of ipv4 e.g. ip_1=192, ip_2 = 168, ip_3=202, ip_4=24
 - then each of those is converted to 8-bit binary and the 32 digits are then concatenated
 - the same is done for ipv6 using 16-bit binary

```{r}
IPBinary <- function(vector=test_ip[,'srcIP'],vec_ind=src_ind) {
  t<-data.table(vector)
  
  colnames(t)<-'t'
  
  #data.table made up of ipv6 addresses and corresponding index in our initial dataset
  t1<-data.table('ip'=t[vec_ind],'index'=vec_ind)
  #data.table made up of ipv4 addresses and corresponding index in our initial dataset
  t2<-data.table('ip'=t[-vec_ind],'index'=(1:L)[-vec_ind])
  
  #separate ipv6 into 8 columns
  t1_sep<-t1%>% separate(ip.t,c(as.character(1:8)))
  #separate ipv4 into 4 columns
  t2_sep<-t2%>% separate(ip.t,c(as.character(1:4)))
  
  #get all ip columns as a matrix
  m1 = as.matrix(t1_sep[,-'index'])
  #turn each hexadecimal to binary
  m1[] = R.utils::intToBin(strtoi(m1,base=16L))
  
  t1_comb<-data.table(m1)
  #concat all 16-bit binary columns together
  t1_conc<-col_concat(t1_comb, sep = "")
  #create data.table with the concatenated binaries and corresponding index
  t1_conc <- data.table('bin'=t1_conc, 'index'=t1[,'index'])
  colnames(t1_conc)<-c('bin','index') 
  
  #repeat for ipv4
  m2 = as.matrix(t2_sep[,-'index'])
  m2[] = R.utils::intToBin(strtoi(m2))
  t2_comb<-data.table(m2)
  t2_conc<-col_concat(t2_comb, sep = "")
  t2_conc <- data.table('bin'=t2_conc, 'index'=t2[,'index'])
  colnames(t2_conc)<-c('bin','index') 
  
  #create a dummy vector of length 200,000
  check_vec<-rep(1,L)
  check_vec
  
  #assign binary value to ith element of vector where i corresponds to appropriate index for ipv6
  for (i in seq_along(t1_conc[[1]])){
    #concatenated binary
    bin<-t1_conc[i][[1]]
    #corresponding index
    ind<-t1_conc[i][[2]]
    
    #value assignment
    check_vec[ind]<-bin
    
  }
  
  pb = txtProgressBar(min = 0, max =length(t2_conc[[1]]), initial = 0)
  #repeat for ipv4
  for (i in seq_along(t2_conc[[1]])){
    
    bin<-t2_conc[i][[1]]
    ind<-t2_conc[i][[2]]
  
    check_vec[ind]<-bin
    
    setTxtProgressBar(pb,i)
  }
  
  return (as.numeric(check_vec))
}
```

We now apply the function to get our final_df
```{r}
final_df<-data.table(random_sample1)
final_df<-final_df[,c('srcIP','dstIP'):=NULL]

final_df$srcIP_bin<-IPBinary()
final_df$dstIP_bin<-IPBinary(vector=test_ip[,'dstIP'])

#this column was set as character, convert to numeric for xgboost
final_df <- final_df[,ts:=as.numeric(ts)]

#we also create a column that corresponds to ipv6 observations
is_ipv6 = rep(0,L)
is_ipv6[src_ind] <- 1
final_df$is_IPv6 <- is_ipv6

sapply(final_df,class)
```
Save final dataframe.
```{r}
write.csv(final_df,'../data/processed/labelled_clean.csv')
```
