---
title: "Data Formatting"
output: html_document
---
# Attempts to label all conn.log
This document is the unused code for formatting conn.log. 


### Main approach 1
Here I started by breaking up the unordered conn.log into 10,000 blocks. The idea behind this was that instead of needing to search the whole of conn.log for a matching packet, it only needed to match a subsection of that. For each snort log check for any matching values in conn.log by progressing down the table. First check the two blocks and if no matches are found then check the rest of the table. This allows for a performance increase as it checks the most likely blocks then checks the rest of the table. When checking the rest of the table if it matches a block further down, then it would only look at that block and below for proceeding snort logs. This means that we are able to make use of the little structure the data has to limit the processing. However, this proved to take too long (more than 2 days) to run and so I tried another approach to see if I could speed it up. 
```{r}
conn2 = {}
i=1
j=1
#table 1 and 2 are the 'current' blocks that are being looked at
table1 = conn1[[j]]
table2 = conn1[[j+1]]
#for each snort log - this is used as this has fewer values than conn and the way that the values are assigned in data.table
while(i < nrow(snort)){
  if (snort$Port[i]){
    #checks if there is an entry that matches the snort log in table 1
    if (nrow(table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],]) > 0){
      #checks if the matching conn entries have not already been matched
      if (table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],'malicious'] != TRUE){
        #If they haven't been matched then label the matching logs as malicious
        table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],'malicious'] <- TRUE 
      }else{
        #if the entries have been matched already, it is possible that there are enteries in table2 that might match the snort log
        if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],]) > 0){
          #if a matching entry exists in table2, then assign the enteries as malicious, append table 1 to the labelled conn log and then move the current logs along 1. This can be done because if we have got to this stage, then there are no more snort logs that will match to table 1 entries due to the structure shared by the snort and conn logs. 
          table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],'malicious'] <- TRUE
          conn2 <- bind_rows(conn2,table1)
          j=j+1
          print(j)
          table1 = table2
          # used to ensure we don't try to access an invalid memory at the end of the table.
          if(j<100001){
            table2 = conn1[[j+1]]  
          }
        }
        #progress to the next snort log.
        i=i+1
      }
    }else {
      #if there are no matching values in table1, check table2
      if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],]) > 0){
        #if a matching entry exists in table2, then assign the enteries as malicious, append table 1 to the labelled conn log and then move the current logs along 1. This can be done because if we have got to this stage, then there are no more snort logs that will match to table 1 entries due to the structure shared by the snort and conn logs. 
        table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],'malicious'] <- TRUE
        conn2 <- bind_rows(conn2,table1)
        j=j+1
        print(j)
        table1 = table2
        # used to ensure we don't try to access an invalid memory at the end of the table.
        if(j<100001){
          table2 = conn1[[j+1]]  
        }
        #progress to the next snort log.
        i=i+1
      }else{
        #if it is neither in table1 or table2, then iterate through the rest of the table until you get to the end or you find a matching value
        k=j+2
        while (nrow(conn1[[k]][conn1[[k]]$ts == snort$ts[i] & conn1[[k]]$srcIP == snort$srcIP[i] & conn1[[k]]$dstIP == snort$dstIP[i] & conn1[[k]]$srcPort == snort$srcPort[i] & conn1[[k]]$dstPort == snort$dstPort[i],]) == 0 & k <= 10000){
          k=k+1
        }
        if (nrow(conn1[[k]][conn1[[k]]$ts == snort$ts[i] & conn1[[k]]$srcIP == snort$srcIP[i] & conn1[[k]]$dstIP == snort$dstIP[i] & conn1[[k]]$srcPort == snort$srcPort[i] & conn1[[k]]$dstPort == snort$dstPort[i],]) > 0){
          #If a matching value is found, then append the previous blocks to the labelled conn.log as there are no more snort logs that will match to those block. 
          conn2 <- bind_rows(conn2,table1)
          for (l in c(j+1:k-1)){
            conn2 <-  bind_rows(conn2,conn1[[l]]) 
          }
          j=k
          print(j)
          #move the 'current' blocks to the block where this matching value was found
          table1 = conn1[[j]]
          if(j<100001){
            table2 = conn1[[j+1]]  
          }
          #assign the matching value as malicious
          table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],'malicious'] <- TRUE
        }
      }
      #progress to the next snort log.
      i=i+1
    }
  }else{
    #If there are no ports in the snort log then you have to rely only on the ts, and src/dst IPs for matching. Beyond this, the code is the same as if there was ports.
    if (nrow(table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],]) > 0){
      if (table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],'malicious'] != TRUE){
        table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],'malicious'] <- TRUE 
      }else{
        if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],]) > 0){
          table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],'malicious'] <- TRUE
          conn2 <- bind_rows(conn2,table1)
          j=j+1
          print(j)
          table1 = table2
          if(j<100001){
            table2 = conn1[[j+1]]  
          }
        }
        i=i+1
      }
    }else {
      if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],]) > 0){
        table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],'malicious'] <- TRUE
        conn2 <- bind_rows(conn2,table1)
        j=j+1
        print(j)
        table1 = table2
        if(j<100001){
          table2 = conn1[[j+1]]  
        }
        i=i+1
      }else{
        k=j+2
        while (nrow(conn1[[k]][conn1[[k]]$ts == snort$ts[i] & conn1[[k]]$srcIP == snort$srcIP[i] & conn1[[k]]$dstIP == snort$dstIP[i],]) == 0 & k <= 10000){
          k=k+1
        }
        if (nrow(conn1[[k]][conn1[[k]]$ts == snort$ts[i] & conn1[[k]]$srcIP == snort$srcIP[i] & conn1[[k]]$dstIP == snort$dstIP[i],]) > 0){
          conn2 <- bind_rows(conn2,table1)
          for (l in c(j+1:k-1)){
            conn2 <-  bind_rows(conn2,conn1[[l]]) 
          }
          j=k
          print(j)
          table1 = conn1[[j]]
          if(j<100001){
            table2 = conn1[[j+1]]  
          }
          table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],'malicious'] <- TRUE
        }
      }
      i=i+1
    }
  }
}
```

### Main approach 2
After realising that approach 1 took too much time, I decided a similar approach but using bigger blocks instead and only checking 2 and not checking the rest of the table when there is no matches in either table. I could have chosen smaller blocks but I found that there was a gap large enough in the snort log data that meant if I did any smaller blocks then this approach would not have worked. After running this code for about a day and not getting past the first block, this approach was slower than the previous approch. As such, we decided to only look at a subset of the data. 

```{r}
x = nrow(conn) %/% 5
i=1
j=1
#table 1 and 2 are the 'current' blocks that are being looked at
table1 = conn[((j-1)*x+1):(j*x+1),]
table2 = conn[((j)*x+1):((j+1)*x+1),]
#next is an indicator to tell the program to progress the current blocks
Next = FALSE
#for each snort log - this is used as this has fewer values than conn and the way that the values are assigned in data.table
while(i < nrow(snort)){
  Next=FALSE
  if (snort$Port[i]){
    #checks if there is an entry that matches the snort log in table 1
    if (nrow(table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],]) > 0){
      #checks if the matching conn entries have not already been matched
      if (table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],'malicious'] != TRUE){
        #If they haven't been matched then label the matching logs as malicious
        table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i] & table1$srcPort == snort$srcPort[i] & table1$dstPort == snort$dstPort[i],'malicious'] <- TRUE 
      }else{
        #if the entries have been matched already, it is possible that there are enteries in table2 that might match the snort log
        if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],]) > 0){
          #if a matching entry exists in table2, then assign the entries as malicious, and indicate that we need to move 'current' blocks
        table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],'malicious'] <- TRUE
        Next <- TRUE
        }
        #progress to the next snort log.
        i=i+1
      }
    }else {
      #if there are no matching values in table1, check table2
      if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],]) > 0){
         #if a matching entry exists in table2, then assign the entries as malicious, and indicate that we need to move 'current' blocks
        table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i] & table2$srcPort == snort$srcPort[i] & table2$dstPort == snort$dstPort[i],'malicious'] <- TRUE
        Next <- TRUE
      }
      #progress to the next snort log.
      i=i+1
    }
  }else{
    #If there are no ports in the snort log then you have to rely only on the ts, and src/dst IPs for matching. Beyond this, the code is the same as if there was ports.
    if (nrow(table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],]) > 0){
      if (table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],'malicious'] != TRUE){
        table1[table1$ts == snort$ts[i] & table1$srcIP == snort$srcIP[i] & table1$dstIP == snort$dstIP[i],'malicious'] <- TRUE 
      }else{
        if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],]) > 0){
          table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],'malicious'] <- TRUE
          Next <- TRUE
        }
        i=i+1
      }
    }else {
      if (nrow(table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],]) > 0){
        table2[table2$ts == snort$ts[i] & table2$srcIP == snort$srcIP[i] & table2$dstIP == snort$dstIP[i],'malicious'] <- TRUE
        i=i+1
        Next <- TRUE
      }else{
        i=i+1
      } 
    }
  }
  #If the 'current' blocks need to be progressed then:
  if (Next){
    print(j)
    table1 <- table2
    if (j == 1){
      malconn <- table1
      table2 = conn[((j+1)*x+1):((j+2)*x+1),]
    }else{
      if (j < 3) { 
        table2 = conn[((j+1)*x+1):((j+2)*x+1),]
      }else if (j == 3){
        table2 = conn[((4)*x+1):nrow(conn),]
      }
      malconn <- rbind(malconn,table1)
    }
    j=j+1
  }
}
```


