---
title: "01-Introduction and Data Formatting"
output: html_document
---

## The brief

The brief was to:

* Choose a goal for our analysis.
* choose a baseline model to compare our complex model to.
* Stratergise how to optimise the parameters of our complex model
* Use a strategy to analyse performance
* Compare the performance of the baseline to the complex model

We decided that we would stick with a simple goal of classifying if a packet was malicious or not as we looked to do more work in other aspects of the project. We decided to use a random forest model as our baseline model because xgboost works in a similar way to random forests but generally performs better so we wanted to look at how much better it performed when the xgboost has tuned hyper parameters and a baseline random forest is used. We intended to use a 10-fold cross validation for the baseline model, however the model was taking too long to run and so we had to use a simple 70/30% training/testing split. We decided to use an ROC curve and AUC to analyse our models as this would give us to see if there were any points where the random forest was better than the xgboost. We also wanted to look more into how xgboost works and some ways that we can optimise its implementation.

## Library requirements

```{r}
if (!require("R.utils")) install.packages("R.utils")
library("R.utils")
if (!require("data.table")) install.packages("data.table")
library("data.table")
if (!require("googledrive")) install.packages("googledrive")
library("googledrive")
```


## The data

We each started by searching out a dataset of connections along with a snort log of another dataset, of which the connections log was a subset. 

To obtain this dataset in a convenient format, we will download it and, if necessary, process it into a standard form. Throughout various parts of our project we checkpointed our progress by saving files to the processed folder.

We place data in the raw or processed folder depending on the stage of processing, both of which are in the `data` folder of our root. So our file system will look like this:

* /data
  * /data/processed
    * /data/processed/snort.csv
    * /data/processed/labelled.csv
    * /data/processed/labelled1.csv
    * /data/processed/labelled_clean.csv
  * /data/raw
    * /data/raw/conn.log

### Create Folders
```{r}
if (!dir.exists('../data')) dir.create('../data')
if (!dir.exists('../data/raw')) dir.create('../data/raw')
if (!dir.exists('../data/processed')) dir.create('../data/processed')
```

### Download the data
```{r}
# if (!file.exists('../data/processed/conn.log'))
if (!file.exists('../data/raw/conn.log.gz') && !file.exists('../data/raw/conn.log')) {
  download.file('https://www.secrepo.com/maccdc2012/conn.log.gz', '../data/raw/conn.log.gz')
}
if (!file.exists('../data/raw/conn.log')) gunzip('../data/raw/conn.log.gz')
```

```{r}
if (!file.exists('../data/processed/snort.csv')) {
  drive_deauth()
  drive_download(as_id("1yx3enHgjRS234E6giO-88qSMoGiJndBo"), path = "../data/processed/snort.csv", overwrite = TRUE)
}
``` 
